---
title: "MTH422A_assigment_3_Solution"
author: "Vijay Soren (211163)"
date: "2024-03-11"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1
### (a)
```{r}
County <- 1:10
Approve <- c(12,90,80,5,63,15,67,22,56,33)
Disapprove <- c(50,150,63,10,63,8,56,19,63,19)
data <- data.frame(County, Approve,Disapprove)
samples_pro <- data$Approve/(data$Approve + data$Disapprove)
mean_samples_pro <- mean(samples_pro)
var_samples_pro <- var(samples_pro)

###a
Y <- data$Approve
n_is <- data$Approve + data$Disapprove
L <- qbeta(c(0.025),shape1 = Y + 0.5, shape2 = n_is - Y + 0.5)
U <- qbeta(c(0.975),shape1 = Y + 0.5, shape2 = n_is - Y + 0.5)
CI95 <- cbind(L,U)
CI95_data <- data.frame(data$County,CI95)
print("95% credibles interval")
print(CI95_data)

```


### (b)
I have found $a = 5.5104$ and $b = 5.9696$


### (c)
```{r}
c <- 0.480
c_star <- c/(1-c)

d <- (c_star + 1)^2
e <- 0.02/c_star
f <- 1/(e*d) - 1
b <- f/(c_star + 1)
a <- c_star*b



###part c
Y <- data$Approve
n_is <- data$Approve + data$Disapprove
L <- qbeta(c(0.025),shape1 = Y + a, shape2 = n_is - Y + b)
U <- qbeta(c(0.975),shape1 = Y + a, shape2 = n_is - Y + b)
CI95_EB <- cbind(L,U)
CI95_data_EB <- data.frame(data$County,CI95_EB)
print("95% credibles interval when beta prior")
print(CI95_data_EB)

```


### (d)
both part results from (a) and (d)  quite similar and all samples proportion's lies within their  95% credibles interval


# 2


### (a)
$\mu_{map}$ = $\sum_{i=1}^{n}\frac{Y_i}{\sigma_i^2}/\sum_{i=1}^{n}\frac{1}{\sigma_i^2}$


### (b)
```{r }
n <- 3

Map_est <- function(Y,sigma_s)
{
  upper <- sum(Y/sigma_s^2)
  lower <- sum(1/sigma_s^2)
  mu_hat <- upper/lower
  return(mu_hat)
  
}


###part b
n = 3
Y <- c(12,10,22)   # y1 = 12, y2 = 10, y3 = 22

sigma_s <- c(3,3,10) 
mu_hat_map <- Map_est(Y = Y, sigma_s = sigma_s)
print((mu_hat_map))
```



### (c)
```{r}
set.seed(2)
n <- 1000
mu <- 5
sigma.s <- sample(1:40,n, replace = TRUE) ## random genration of sigma_s bw 1 :40 of length n
Y <- rnorm(n, mean = mu, sd = sigma.s)
mu_samples_pos <- rnorm(n, mean =  sum(Y/sigma.s^2)/(sum(1/sigma.s^2)), sd = sqrt(1/sum(1/sigma.s^2)))
 ## posterior mean
mean(mu_samples_pos)
print("mean of posterior")
print(mean(mu_samples_pos))

```




### (d)
```{r}
set.seed(2)
n <- 1000
mu <- 5
sigma.s <- sample(1:40,n, replace = TRUE) ## random genration of sigma_s bw 1 :40 of length n
Y <- rnorm(n, mean = mu, sd = sigma.s)

## posterior distribution
mu.pos.var <- 1/(sum(1/sigma.s^2))
mu.pos.mean <- sum(Y/sigma.s^2)/sum(1/sigma.s^2)

MAP_mu <- sum(Y/sigma.s^2)/sum(1/sigma.s^2)

mu.pos.samples <- rnorm(1e4, mean = mu.pos.mean, sd = sqrt(mu.pos.var))
mu.pos <- sort(mu.pos.samples)
density <- dnorm(mu.pos, mean = mu.pos.mean, sd = sqrt(mu.pos.var))
plot(mu.pos, density, type = "l")
abline(v = MAP_mu , col = "red", lty = 2)
abline(v = mu.pos.mean , col = "green", lty = 6)
legend("topright", c("MAP_est","posterior mean"), fill = c("red","green"), lty = 2)
```



# 3


### (a)
The full conditional posterior distributions for $\sigma_1^2$ is $Inverse-Gamma(a + 0.5,Y_1^2/2 +b)$


The full conditional posterior distributions for $b$ is $Gamma(1,1+\sum_{i=1}^{n}\frac{1}{\sigma_i^2})$


### (b) psudocode

```{r, echo=T}
set.seed(1)
library(invgamma)
n <- 10
a <- 10
b <- rgamma(1,shape = 1,rate = 1)
sigmaSq <- rinvgamma(n,shape = a, rate = b)

Y <- rnorm(n, mean = 0, sd = sqrt(sigmaSq))



b.update  <- function(sigmaSq)
{
  b.pos <- rgamma(1,shape = 1, rate = 1 + sum(1/sigmaSq))
  return(b.pos)
}

sigmaSq.update <- function(Y, a, b,n)
{
  sigmaSq<- rinvgamma(n, shape = a + 1/2, rate = b + 0.5*sum(Y^2))
  
  return(sigmaSq)
}

MCMC <- function(Y, b.init, sigmaSq.init, iters){
  
  # chain initiation
  b <- b.init
  n <- length(Y)
  sigmaSq <- rep(sigmaSq.init, n)
  
  # define chains
  b.chain <- rep(NA, iters)
  sigmaSq.chain <- matrix(NA, iters, n)
  
  # start MCMC
  for(i in 1:iters){
    sigmaSq <- sigmaSq.update(Y, a, b,n)
    b <- b.update(sigmaSq)
    
    
    b.chain[i] <- b
    sigmaSq.chain[i,] <- sigmaSq
    
  }
  
  # return chains
  out <- list(b.chain = b.chain, 
              sigmaSq.chain = sigmaSq.chain)
  return(out)}
```



### (c) 
### a = 10
```{r}
set.seed(1)
library(invgamma)
n <- 10
a <- 10
b <- rgamma(1,shape = 1,rate = 1)
sigmaSq <- rinvgamma(n,shape = a, rate = b)

Y <- rnorm(n, mean = 0, sd = sqrt(sigmaSq))



b.update  <- function(sigmaSq)
{
  b.pos <- rgamma(1,shape = 1, rate = 1 + sum(1/sigmaSq))
  return(b.pos)
}

sigmaSq.update <- function(Y, a, b,n)
{
  sigmaSq<- rinvgamma(n, shape = a + 1/2, rate = b + 0.5*sum(Y^2))
  
  return(sigmaSq)
}

MCMC <- function(Y, b.init, sigmaSq.init, iters){
  
  # chain initiation
  b <- b.init
  n <- length(Y)
  sigmaSq <- rep(sigmaSq.init, n)
  
  # define chains
  b.chain <- rep(NA, iters)
  sigmaSq.chain <- matrix(NA, iters, n)
  
  # start MCMC
  for(i in 1:iters){
    sigmaSq <- sigmaSq.update(Y, a, b,n)
    b <- b.update(sigmaSq)
    
    
    b.chain[i] <- b
    sigmaSq.chain[i,] <- sigmaSq
    
  }
  
  # return chains
  out <- list(b.chain = b.chain, 
              sigmaSq.chain = sigmaSq.chain)
  return(out)}


MCMC.out <- MCMC(Y = Y,
                 b.init = 0, 
                 sigmaSq.init = .04,
                 iters = 30000)


# marginal plot
plot(density(MCMC.out$b.chain), xlab = expression(b))
plot(density(MCMC.out$sigmaSq.chain[,1]),xlab = "sigmaSq1")
plot(density(MCMC.out$sigmaSq.chain[,2]),xlab = "sigmaSq2")
plot(density(MCMC.out$sigmaSq.chain[,3]),xlab = "sigmaSq3")
plot(density(MCMC.out$sigmaSq.chain[,4]),xlab = "sigmaSq4")
plot(density(MCMC.out$sigmaSq.chain[,5]),xlab = "sigmaSq5")
plot(density(MCMC.out$sigmaSq.chain[,6]),xlab = "sigmaSq6")
plot(density(MCMC.out$sigmaSq.chain[,7]),xlab = "sigmaSq7")
plot(density(MCMC.out$sigmaSq.chain[,8]),xlab = "sigmaSq8")
plot(density(MCMC.out$sigmaSq.chain[,9]),xlab = "sigmaSq9")
plot(density(MCMC.out$sigmaSq.chain[,10]),xlab = "sigmaSq10")


```


### (d) 
### a =1
```{r}
set.seed(1)
library(invgamma)
n <- 10
a <- 1
b <- rgamma(1,shape = 1,rate = 1)
sigmaSq <- rinvgamma(n,shape = a, rate = b)

Y <- rnorm(n, mean = 0, sd = sqrt(sigmaSq))



b.update  <- function(sigmaSq)
{
  b.pos <- rgamma(1,shape = 1, rate = 1 + sum(1/sigmaSq))
  return(b.pos)
}

sigmaSq.update <- function(Y, a, b,n)
{
  sigmaSq<- rinvgamma(n, shape = a + 1/2, rate = b + 0.5*sum(Y^2))
  
  return(sigmaSq)
}

MCMC <- function(Y, b.init, sigmaSq.init, iters){
  
  # chain initiation
  b <- b.init
  n <- length(Y)
  sigmaSq <- rep(sigmaSq.init, n)
  
  # define chains
  b.chain <- rep(NA, iters)
  sigmaSq.chain <- matrix(NA, iters, n)
  
  # start MCMC
  for(i in 1:iters){
    sigmaSq <- sigmaSq.update(Y, a, b,n)
    b <- b.update(sigmaSq)
    
    
    b.chain[i] <- b
    sigmaSq.chain[i,] <- sigmaSq
    
  }
  
  # return chains
  out <- list(b.chain = b.chain, 
              sigmaSq.chain = sigmaSq.chain)
  return(out)}


MCMC.out <- MCMC(Y = Y,
                 b.init = 0, 
                 sigmaSq.init = .04,
                 iters = 30000)


# marginal plot
plot(density(MCMC.out$b.chain), xlab = expression(b))
plot(density(MCMC.out$sigmaSq.chain[,1]),xlab = "sigmaSq1")
plot(density(MCMC.out$sigmaSq.chain[,2]),xlab = "sigmaSq2")
plot(density(MCMC.out$sigmaSq.chain[,3]),xlab = "sigmaSq3")
plot(density(MCMC.out$sigmaSq.chain[,4]),xlab = "sigmaSq4")
plot(density(MCMC.out$sigmaSq.chain[,5]),xlab = "sigmaSq5")
plot(density(MCMC.out$sigmaSq.chain[,6]),xlab = "sigmaSq6")
plot(density(MCMC.out$sigmaSq.chain[,7]),xlab = "sigmaSq7")
plot(density(MCMC.out$sigmaSq.chain[,8]),xlab = "sigmaSq8")
plot(density(MCMC.out$sigmaSq.chain[,9]),xlab = "sigmaSq9")
plot(density(MCMC.out$sigmaSq.chain[,10]),xlab = "sigmaSq10")
```



### (e)
### jags part
```{r}

library(rjags)
## a=10
a <- 10

data <- list(Y = Y, n = length(Y), a = a)
model_string <- textConnection("model{
# Likelihood (dnorm uses a precision, not variance)
for(i in 1:n){
Y[i] ~ dnorm(0,tau[i]) #tau = 1/sigma.sq
}

#prior
for(i in 1:n){
  tau[i] ~ dgamma(a, b)
  sigma.sq[i] <- 1/tau[i]
}

b ~ dgamma(1,1)

}")
tau_init <- rep(0.04, n)
inits <- list(tau = tau_init, b = 0.02)
model <- jags.model(model_string, data = data, inits = inits, quiet = TRUE)
update(model, 10000, progress.bar = "none")
params <- c("b", "sigma.sq")
samples <- coda.samples(model,
                        variable.names = params,
                        n.iter = 20000, progress.bar = "none")

summary(samples)


plot(samples)

```



both have similar results



# 4

### (a)
because when calculate mean of $\theta_i$ prior then we get $q_i$ which is belong to $(0,1)$


### (b)


### (c)
The full conditional posterior for $\theta_1$ is $Beta(Y_1 + q_1*exp(m),  n_1 - Y_1 + (1-q_i)*exp(m))$


### (d)
```{r}
proportion <- c(0.845,0.847,0.880,0.674,0.909,0.898,0.770,0.801,0.802,0.875)
clucth_make <- c(64,72,55,27,75,24,28,66,40,13)
clucth_attm <- c(75,95,63,39,83,26,41,82,54,16)

player_nam <- c("Russell Westbrook ","James Harden ","Kawhi Leonard ","LeBron James",
                "Isaiah Thomas","Stephen Curry","Giannis Antetokounmpo",
                "John Wall","Anthony Davis","Kevin Durant")
data <- data.frame(player_nam,proportion,clucth_make,clucth_attm)
n <- 10
Y<- data$clucth_make
theta <- data$proportion
n_s <- data$clucth_attm
q <- runif(n) 
m.update <- function(m,theta, q, sdev =4)
{
  m_j <- m
  m_c <- rnorm(1,mean = m,sd = sdev)
  Low <- (m^2)/20
  Upp <- (m_c^2)/20
  for (i in 1:n) {
    Low <- Low + (q[i]* exp(m) - 1)*log(theta[i]) + ((1- q[i])*exp(m) -1)*log(1-theta[i])
  }
  
  for (i in 1:n) {
    Upp <- Upp + (q[i]* exp(m_c) - 1)*log(theta[i]) + ((1- q[i])*exp(m_c) -1 )*log(1-theta[i])
  }
  logR <- Upp - Low
  if(logR == "NaN")
  {
    return(m_j)
  }
  U <- runif(1) # U(0,1)
  if(log(U) < logR){
    m_j <- m_c
  }
  
  return(m_j)
}


theta.update <- function(Y,q,n,m, n_s)
{
  theta.pos <- rbeta(n, shape1 = Y + q*exp(m), shape2 = n_s - Y +exp(1- q))
  return(theta.pos)
}


MCMC <- function(Y, m.init, theta.init, a, b, iters){
  
  # chain initiation
  m <- m.init
  n <- length(Y)
  theta <- rep(theta.init, n)
  
  # define chains
  m.chain <- rep(NA, iters)
  theta.chain <- matrix(NA, iters, n)
  
  # start MCMC
  for(i in 1:iters){
    theta <- theta.update(Y,q,n, m,n_s )
    m <- m.update(m,theta , q)
    
    
    m.chain[i] <- m
    theta.chain[i,] <- theta
    
  }
  
  # return chains
  out <- list(m.chain = m.chain, 
              theta.chain = theta.chain)
  return(out)}

# RUN MCMC

MCMC.out <- MCMC(Y = Y,
                 m.init = 1, 
                 theta.init = 0.4,
                 iters = 30000,n)


theta.pos.samples <- MCMC.out$theta.chain
theta1_95 <- quantile(theta.pos.samples[,1], c(0.025,0.975))
theta2_95 <- quantile(theta.pos.samples[,2], c(0.025,0.975))
theta3_95 <- quantile(theta.pos.samples[,3], c(0.025,0.975))
theta4_95 <- quantile(theta.pos.samples[,4], c(0.025,0.975))
theta5_95 <- quantile(theta.pos.samples[,5], c(0.025,0.975))
theta6_95 <- quantile(theta.pos.samples[,6], c(0.025,0.975))
theta7_95 <- quantile(theta.pos.samples[,7], c(0.025,0.975))
theta8_95 <- quantile(theta.pos.samples[,8], c(0.025,0.975))
theta9_95 <- quantile(theta.pos.samples[,9], c(0.025,0.975))
theta10_95 <- quantile(theta.pos.samples[,10], c(0.025,0.975))


m.pos.samples <- MCMC.out$m.chain
m.pos_95 <- quantile(m.pos.samples, c(0.025,0.975))

ci95 <- rbind(m.pos_95 ,theta1_95, theta2_95, theta3_95,
              theta4_95, theta5_95, theta6_95,theta7_95,
              theta8_95, theta9_95,theta10_95)
print("95% credibles interval")
print(ci95)

```


### (e)
```{r}
proportion <- c(0.845,0.847,0.880,0.674,0.909,0.898,0.770,0.801,0.802,0.875)
clucth_make <- c(64,72,55,27,75,24,28,66,40,13)
clucth_attm <- c(75,95,63,39,83,26,41,82,54,16)

player_nam <- c("Russell Westbrook ","James Harden ","Kawhi Leonard ","LeBron James",
                "Isaiah Thomas","Stephen Curry","Giannis Antetokounmpo",
                "John Wall","Anthony Davis","Kevin Durant")
data <- data.frame(player_nam,proportion,clucth_make,clucth_attm)
n <- 10
Y<- data$clucth_make
theta <- data$proportion
n_s <- data$clucth_attm
q <- runif(n)
data <- list(Y = Y, n = n, q=q, n_s = n_s)
library(rjags)
model_string <- textConnection("model{
# Likelihood 
for(i in 1:n){
Y[i] ~ dbin(theta[i],n_s[i])
}

#prior
for(i in 1:n){
theta[i] ~ dbeta(q[i] * exp(m),exp(m)*(1-q[i]) )
}
m ~ dnorm(0,0.1)

}")
theta_inti <- rep(0.4,n)
inits <- list( theta = theta_inti,m = 3.3)
model <- jags.model(model_string, data = data, inits = inits, quiet = TRUE)
update(model, 10000, progress.bar = "none")
params <- c("m", "theta")
samples <- coda.samples(model,
                        variable.names = params,
                        n.iter = 20000, progress.bar = "none")

summary(samples)
plot(samples)

```
 
 
 yes both code have very close answer of 95% credibles interval


 
### (f)
JAGS are easy use here we don't have write all marginal posterior distribution


but writing own MCMC code we have find their marginal posterior distribution


# 8
```{r,echo =TRUE}
library(MASS)
data(galaxies)
Y <- galaxies


mu.update <- function(Y, mu, sigma)
{
  mu_j <- mu
  mu_c <- rnorm(1, mean = mu, sd = 2)
  R <- exp(sum(abs(Y - mu)) - sum(abs(Y - mu_c)))
  R <- min(1,R)
  U <- runif(1)
  if(U < R)
  {
    mu_j <- mu_c
  }
  return(mu_j)
}
sigma.update <- function(Y, mu, sigma)
{
  n <- length(Y)
  sigma_j <- sigma
  sigma_c <- rnorm(1, mean = sigma, sd = 2)
  R <- (sigma/(sigma_c))^n * exp(sum(Y -mu)*(1/sigma - 1/sigma_j))
  R <- min(1,R)
  if(runif(1) < R)
  {
    sigma_j <- sigma_c
  }
  return(sigma_j)
}

MCMC <- function(Y, mu.init, sigma.init, iters){
  
  # chain initiation
  mu <- mu.init
  n <- length(Y)
  sigma <- sigma.init
  
  # define chains
  mu.chain <- rep(NA, iters)
  sigma.chain <-rep(NA, iters)
  # start MCMC
  for(i in 1:iters){
    sigma <- sigma.update(Y, mu, sigma )
    mu <- mu.update(Y, mu, sigma)
    
    
    mu.chain[i] <- mu
    sigma.chain[i] <- sigma
    
  }
  
  # return chains
  out <- list(mu.chain = mu.chain, 
              sigma.chain = sigma.chain)
  return(out)}

# RUN MCMC

MCMC.out <- MCMC(Y = Y,
                 mu.init = 1, 
                 sigma.init = 0.4,
                 iters = 30000)


# marginal plot
plot(density(MCMC.out$mu.chain))
plot(density(MCMC.out$sigma.chain))
```